### 2018-08-09 paper/abstract.tex
Use acro package to handle citation of Tor

% -*- tex-main-file: "contents.tex" compile-command: "pdflatex p2p-private-cloud.tex" ispell-dictionary: "american" -*-
% What's the problem?
% Why is it a problem? Research gap left by other approaches?
% Why is it important? Why care?
% What's the approach? How to solve the problem?
% What's the findings? How was it evaluated, what are the results, limitations, 
% what remains to be done?
%
%abstract
%problem/topic
This paper proposes SP$^2$ORES, a novel private file transfer service that avoids any third-party storage or centralized bootstrapping service.
%research gap
Our solution, which adopts a fully decentralized design, overcomes the limitations of cloud file-sharing systems such as Dropbox, which require trust in the service provider and can be monitored, even when their data is encrypted. Our proposal also goes beyond existing private peer-to-peer (P2P) solutions that must rely on storage replication for efficiency or a fixed set of trusted servers for bootstrapping purposes, thus limiting their level of protection. % We are interested in minimizing the trust needed in other systems and preserving the users' privacy in terms of data and metadata, what they share with whom and when.
%why now
% Nowadays, many people use multiple devices and it has thus become possible to benefit from P2P technology for one's own devices alone, thereby reducing one's exposure to security and privacy threats from third parties.
%approach
In this paper, we demonstrate that users do not necessarily have to choose between efficiency and strong privacy when sharing data with others. % that is  both efficient and strongly private is possible.
\commentFT{Trying to highlight POR, as likely to catch people's attention, and contribute to the paper's long-term interest.} At the core of our proposal lies a novel stateless probabilistic onion routing protocol called SP$^2$OR, which exploits decentralized peer sampling, a probabilistic model of node availability, and a novel set-based onion routing header format (dubbed SphinxES). \commentFT{Sharing node availability information freely could potentially disclose private information. We might want to discuss this risk somewhere in the paper.}
% Our proposal combines onion routing with probabilistic node selection, informed by peer sampling and Hidden Markov Model predictions for node availability.
%findings/contributions:
The resulting solution is fully decentralized, provides sender-receiver anonymity, and can deliver competitive performance in spite of unreliable nodes. %% to protect user privacy in decentralized file sharing.
% We show that even a relatively naive prediction model suffices to share files despite the unreliable availability of individual nodes. %% The same model can potentially be applied for better-informed churn handling in other p2p systems. In the case of onion routing systems for anonymity,
\commentFT{Using transient to avoid repeating unreliable.}
This design makes it possible to exploit transient nodes, without dedicated relays, %take advantage of locality, 
neither of which is currently the case in Tor.
\sonja{say something about security requirements? About sphinxes?} \commentFT{Added a mention to SphinxES, not sure if we can discuss security requirements without making the abstract overly long.}

>>>>

% -*- tex-main-file: "contents.tex" compile-command: "pdflatex spores.tex" ispell-dictionary: "american" -*-
% What's the problem?
% Why is it a problem? Research gap left by other approaches?
% Why is it important? Why care?
% What's the approach? How to solve the problem?
% What's the findings? How was it evaluated, what are the results, limitations, 
% what remains to be done?
%
%abstract
%problem/topic
This paper proposes SP$^2$ORES, a novel private file transfer service that avoids any third-party storage or centralized bootstrapping service.
%research gap
Our solution, which adopts a fully decentralized design, overcomes the limitations of cloud file-sharing systems such as Dropbox, which require trust in the service provider and can be monitored, even when their data is encrypted. Our proposal also goes beyond existing private peer-to-peer (P2P) solutions that must rely on storage replication for efficiency or a fixed set of trusted servers for bootstrapping purposes, thus limiting their level of protection. % We are interested in minimizing the trust needed in other systems and preserving the users' privacy in terms of data and metadata, what they share with whom and when.
%why now
% Nowadays, many people use multiple devices and it has thus become possible to benefit from P2P technology for one's own devices alone, thereby reducing one's exposure to security and privacy threats from third parties.
%approach
In this paper, we demonstrate that users do not necessarily have to choose between efficiency and strong privacy when sharing data with others. % that is  both efficient and strongly private is possible.
\commentFT{Trying to highlight POR, as likely to catch people's attention, and contribute to the paper's long-term interest.} At the core of our proposal lies a novel stateless probabilistic onion routing protocol called SP$^2$OR, which exploits decentralized peer sampling, a probabilistic model of node availability, and a novel set-based onion routing header format (dubbed SphinxES). \commentFT{Sharing node availability information freely could potentially disclose private information. We might want to discuss this risk somewhere in the paper.}
% Our proposal combines onion routing with probabilistic node selection, informed by peer sampling and Hidden Markov Model predictions for node availability.
%findings/contributions:
The resulting solution is fully decentralized, provides sender-receiver anonymity, and can deliver competitive performance in spite of unreliable nodes. %% to protect user privacy in decentralized file sharing.
% We show that even a relatively naive prediction model suffices to share files despite the unreliable availability of individual nodes. %% The same model can potentially be applied for better-informed churn handling in other p2p systems. In the case of onion routing systems for anonymity,
\commentFT{Using transient to avoid repeating unreliable.}
This design makes it possible to exploit transient nodes, without dedicated relays, %take advantage of locality, 
neither of which is currently the case in \ac{Tor}.
\sonja{say something about security requirements? About sphinxes?} \commentFT{Added a mention to SphinxES, not sure if we can discuss security requirements without making the abstract overly long.}

### 2018-08-09 paper/approach.tex
Use acro package to handle citation of Tor

\Sphinxes does \emph{not} introduce any restrictions on whether the network 
topology must be fully connected or not, be asynchronous (as in Tor) or 
synchronous (as in mix-nets), whether node selection must be uniformly random 
or not.
Just as Sphinx, it can be used in many different settings.

>>>>

\Sphinxes does \emph{not} introduce any restrictions on whether the network 
topology must be fully connected or not, be asynchronous (as in \ac{Tor}) or 
synchronous (as in mix-nets), whether node selection must be uniformly random 
or not.
Just as Sphinx, it can be used in many different settings.

### 2018-08-09 paper/intro.tex
Use acro package to handle citation of Tor

% -*- tex-main-file: "contents.tex" compile-command: "pdflatex p2p-private-cloud.tex" ispell-dictionary: "american" -*-
%!TEX root = p2p-private-cloud.tex

@@

Constructing a system that provides fully anonymous content sharing while avoiding the above pitfalls is unfortunately highly
challenging. The last two decades have witnessed many attempts at building privacy-preserving and anonymous P2P data-sharing
networks~\cite{Clarke:2001,Gnunet2002,Freedman:2002,Nambiar:2006,Rennhard:2002}.
These solutions typically leverage \emph{onion
 routing}~\cite{Chaum:1981}, with some improvements to perform node
 discovery in a distributed manner, a step that is otherwise
centralized in traditional onion routing schemes. However, because of the high
churn typically experienced by P2P networks, the performance of these systems has been
discouraging~\cite{LeBlond:2013}, as churn causes routes to disappear and be
reconstructed frequently, which is particularly costly. By contrast, Tor
\cite{Tor}, one of the most popular anonymous networks,
trades lower anonymity guarantees for a much better
performance. \sonja{lower than who?} Tor's lower protection arises from
a partially centralized design that hinges on a set of \emph{directory
servers}, which are used to advertise trusted nodes that are available to bootstrap onion
routes.

>>>>

% -*- tex-main-file: "contents.tex" compile-command: "pdflatex spores.tex" ispell-dictionary: "american" -*-
%!TEX root = spores.tex

@@

Constructing a system that provides fully anonymous content sharing while avoiding the above pitfalls is unfortunately highly
challenging. The last two decades have witnessed many attempts at building privacy-preserving and anonymous P2P data-sharing
networks~\cite{Clarke:2001,Gnunet2002,Freedman:2002,Nambiar:2006,Rennhard:2002}.
These solutions typically leverage \emph{onion
 routing}~\cite{Chaum:1981}, with some improvements to perform node
 discovery in a distributed manner, a step that is otherwise
centralized in traditional onion routing schemes. However, because of the high
churn typically experienced by P2P networks, the performance of these systems has been
discouraging~\cite{LeBlond:2013}, as churn causes routes to disappear and be
reconstructed frequently, which is particularly costly. By contrast, \ac{Tor}, 
one of the most popular anonymous networks,
trades lower anonymity guarantees for a much better
performance. \sonja{lower than who?} \acg{Tor} lower protection arises from
a partially centralized design that hinges on a set of \emph{directory
servers}, which are used to advertise trusted nodes that are available to bootstrap onion
routes.

### 2018-08-09 paper/security-discussion.tex
Use acro package to handle citation of Tor

A global, passive, network-observing adversary can of course do
statistical disclosure attacks to eventually learn who communicates
with whom~\cite{StatisticalDisclosureAttacks}. There are also denial
of service attacks that can lead to the adversary learning more from
the system~\cite{DenialOfSecurity}.  As mentioned in
\cref{system-model}, we share the adversary model with Tor
%and Sphinx
and thus exclude a global passive adversary. Our focus is on
protecting Alice and Bob from each other and from others who also
participate as nodes in the network, not on protecting them from
surveillance. Nevertheless, we offer some reflections on such an
adversary. We hypothesize that our stateless and redundant design
might yield some advantages over classical \ac{OR}. For
example, no control messages are sent back to the source if the
adversary has introduced an error.  \sonja{``But these advantages will
  only postpone the success of the attack, eventually even our
  protocol must resend.'' why?}  \commentDaniel{Due to statistics. We should 
  also look into
  this denial of security.} Another example is that the source-bounded
random-walk and new routes for file chunks make tracking more
difficult than with regular onion routing circuits. A drawback, however, is
that since the chunks can take several paths from source to destination, the source 
will send and the destination will receive more traffic than other nodes in the 
network~\cite{RoutingSurveyAnonymousProtocols}. However, this is mitigated by the fact that the source and 
destination can be several different devices --- \ie from the \squads.

We would also like to point out that the
vulnerability to a global passive attacker is not inherent in Sphinx
or \Sphinxes, but can be caused by the way they are applied, \ie the
traffic sent. If performance and resource consumption were not an
issue, batched and well-timed message or file transfer by all nodes
would render traffic analysis less efficient. Concretely, \Ac{SPORES} inherits 
the security properties from \Sphinxes and \ac{SPOR}.
However, both \Sphinxes and \ac{SPOR} can be used with any mixing strategy, and 
the security they provide is as secure as the mixing strategy (or flushing 
algorithm) allows.
Since \ac{SPORES} is a fully-fledged system, it must make a concrete choice for 
the mixing strategy.
Since \ac{SPORES} aims at low latency, routers do not batch and mix incoming 
messages; they immediately forward any messages.
This puts \ac{SPORES} in the same position as Tor~\cite{Tor}, \ie more 
vulnerable to traffic analysis than batching 
mix-nets~\cite{RoutingSurveyAnonymousProtocols} \commentDaniel{we should 
  probably look for the original papers} --- this vulnerability is inherited 
from the mixing strategy.

>>>>

A global, passive, network-observing adversary can of course do
statistical disclosure attacks to eventually learn who communicates
with whom~\cite{StatisticalDisclosureAttacks}. There are also denial
of service attacks that can lead to the adversary learning more from
the system~\cite{DenialOfSecurity}.  As mentioned in
\cref{system-model}, we share the adversary model with \ac{Tor}
%and Sphinx
and thus exclude a global passive adversary. Our focus is on
protecting Alice and Bob from each other and from others who also
participate as nodes in the network, not on protecting them from
surveillance. Nevertheless, we offer some reflections on such an
adversary. We hypothesize that our stateless and redundant design
might yield some advantages over classical \ac{OR}. For
example, no control messages are sent back to the source if the
adversary has introduced an error.  \sonja{``But these advantages will
  only postpone the success of the attack, eventually even our
  protocol must resend.'' why?}  \commentDaniel{Due to statistics. We should 
  also look into
  this denial of security.} Another example is that the source-bounded
random-walk and new routes for file chunks make tracking more
difficult than with regular onion routing circuits. A drawback, however, is
that since the chunks can take several paths from source to destination, the source 
will send and the destination will receive more traffic than other nodes in the 
network~\cite{RoutingSurveyAnonymousProtocols}. However, this is mitigated by the fact that the source and 
destination can be several different devices --- \ie from the \squads.

We would also like to point out that the
vulnerability to a global passive attacker is not inherent in Sphinx
or \Sphinxes, but can be caused by the way they are applied, \ie the
traffic sent. If performance and resource consumption were not an
issue, batched and well-timed message or file transfer by all nodes
would render traffic analysis less efficient. Concretely, \Ac{SPORES} inherits 
the security properties from \Sphinxes and \ac{SPOR}.
However, both \Sphinxes and \ac{SPOR} can be used with any mixing strategy, and 
the security they provide is as secure as the mixing strategy (or flushing 
algorithm) allows.
Since \ac{SPORES} is a fully-fledged system, it must make a concrete choice for 
the mixing strategy.
Since \ac{SPORES} aims at low latency, routers do not batch and mix incoming 
messages; they immediately forward any messages.
This puts \ac{SPORES} in the same position as \ac{Tor}, \ie more vulnerable to 
traffic analysis than batching mix-nets~\cite{RoutingSurveyAnonymousProtocols} 
\commentDaniel{we should probably look for the original papers} --- this 
vulnerability is inherited from the mixing strategy.

### 2018-08-09 paper/simpleSPOR.tex
Use acro package to handle citation of Tor

Simple Tor:
1) S: get I from directory
2) set up circuit (via key agreements with each I_i)
3) send/receive in circuit via onion (constructed by encrypting packet for each layer, next hop outermost)

@@

Note: need to figure out how Tor sends both return onion and request, then do the same here.

@@

Comparison between Tor and SPOR:
- SPOR file sharing similar to Tor hidden service
- Tor web browsing similar to Spor
- Tor stateful (circuits), SPOR stateless
- SPOR layers are sets of nodes, Tor layers are single nodes.
- SPOR allows for parallelism (bottlenecks only at S and D). Whereas Tor bottlenecks at the node in the circuit with the lowest 
bandwidth.
- Tor infrastracture, SPOR decentralized
- SPOR allows any node intor relay directory. Anyone can enter. Then they receive different ranks based on "good behaviour". They also monitor the nodes and kick out suspicious nodes.
- SPOR chooses routes informed by availability prediction to minimize churn (prevention)
- SPOR can do local repairs (future work?) to cope with churn (detection, response, recovery). Already now: if node not available, choose another from the set to send to. Problem in current version if node is available but then crashes. [3]
- SPOR can limit location, e.g. within a country? Would help with availability and uniform distribution of chosen nodes again due to lack of a-priori differences in availability in same time zone.
- Tor sender anonymity, SPOR file sharing sender-receiver anonymity (ID of device, not user as there's out-of-band communication), but the more fair analogy is probably with hidden services, where there should be sender-receiver anonymity also in Tor. Only sender anonymity when SPOR is used for web browsing.

@@

[2]
Daniel: We could let the digest be the digest of each chunk. Then we get the ordering of the chunks as well. (\Ie the metadata is equivalent to a torrent file.) Then the ack could be a bitmap of which chunks are received. This is good for efficiency but opens up for attacks. However, the damage of this is not as bad as for mixes (or Tor), since the retransmission will *probably* not take the same route.
Adrien: We take inspiration from BitTorrent. They chunk files in fixed-size
pieces, make a SHA1 hash of each piece, and a hash of the previous
hashes. This is all we need to share. We can even share it in-band,
during your point 4) in file-sharing. Thus, we only need to exchange a
file identifier out of band.

>>>>

Simple \ac{Tor}:
1) S: get I from directory
2) set up circuit (via key agreements with each I_i)
3) send/receive in circuit via onion (constructed by encrypting packet for each layer, next hop outermost)

@@

Note: need to figure out how \ac{Tor} sends both return onion and request, then do the same here.

@@

Comparison between \ac{Tor} and SPOR:
- SPOR file sharing similar to \ac{Tor} hidden service
- \ac{Tor} web browsing similar to Spor
- \ac{Tor} stateful (circuits), SPOR stateless
- SPOR layers are sets of nodes, \ac{Tor} layers are single nodes.
- SPOR allows for parallelism (bottlenecks only at S and D). Whereas \ac{Tor} bottlenecks at the node in the circuit with the lowest 
bandwidth.
- \ac{Tor} infrastracture, SPOR decentralized
- SPOR allows any node intor relay directory. Anyone can enter. Then they receive different ranks based on "good behaviour". They also monitor the nodes and kick out suspicious nodes.
- SPOR chooses routes informed by availability prediction to minimize churn (prevention)
- SPOR can do local repairs (future work?) to cope with churn (detection, response, recovery). Already now: if node not available, choose another from the set to send to. Problem in current version if node is available but then crashes. [3]
- SPOR can limit location, e.g. within a country? Would help with availability and uniform distribution of chosen nodes again due to lack of a-priori differences in availability in same time zone.
- \ac{Tor} sender anonymity, SPOR file sharing sender-receiver anonymity (ID of device, not user as there's out-of-band communication), but the more fair analogy is probably with hidden services, where there should be sender-receiver anonymity also in \ac{Tor}. Only sender anonymity when SPOR is used for web browsing.

@@

[2]
Daniel: We could let the digest be the digest of each chunk. Then we get the ordering of the chunks as well. (\Ie the metadata is equivalent to a torrent file.) Then the ack could be a bitmap of which chunks are received. This is good for efficiency but opens up for attacks. However, the damage of this is not as bad as for mixes (or \ac{Tor}), since the retransmission will *probably* not take the same route.
Adrien: We take inspiration from BitTorrent. They chunk files in fixed-size
pieces, make a SHA1 hash of each piece, and a hash of the previous
hashes. This is all we need to share. We can even share it in-band,
during your point 4) in file-sharing. Thus, we only need to exchange a
file identifier out of band.

### 2018-08-09 paper/system-model.tex
rm error
Use acro package to handle citation of Tor

% -*- tex-main-file: "contents.tex" compile-command: "pdflatex -halt-on-error p2p-private-cloud.tex > /dev/null" ispell-dictionary: "american" -*-
%!TEX root = p2p-private-cloud.tex

@@

%adversary model: remove provider, compensate, results in same as in sphinxOR/Tor?  friends?

@@

In the following, we first describe how file transfer typically occurs
when users rely on a cloud storage provider. We use this simple use
case to highlight some of the desirable privacy properties of a file
transfer service, before mapping them into our fully decentralized
(a.k.a. peer-to-peer, or P2P short short\footnote{We use the terms \emph{fully decentralized} and
  \emph{P2P} interchangeably, and preferentially the latter for brevity.})
context.  We finally consider the additional privacy properties
required to protect users from inferences, such as by network traffic
analysis, and state our adversary model.

@@

\paragraph*{Adversary model} To assess our approach, we use the
adversary model assumed by Tor~\cite{Tor} and Sphinx~\cite{Sphinx}:
the adversary has complete control over some part of the network as
well as some of the nodes. Like them, we do not consider a passive global network
adversary, against which it is notoriously difficult to achieve
security~\cite{SystemsForAnonymousCommunication}.

>>>>

% -*- tex-main-file: "contents.tex" compile-command: "pdflatex -halt-on-error spores.tex > /dev/null" ispell-dictionary: "american" -*-
%!TEX root = spores.tex

@@

%adversary model: remove provider, compensate, results in same as in sphinxOR/\ac{Tor}?  friends?

@@

In the following, we first describe how file transfer typically occurs
when users rely on a cloud storage provider. We use this simple use
case to highlight some of the desirable privacy properties of a file
transfer service, before mapping them into our fully decentralized
(a.k.a. P2P\footnote{We use the terms \emph{fully decentralized} and
  \emph{P2P} interchangeably, and preferentially the latter for brevity.})
context.  We finally consider the additional privacy properties
required to protect users from inferences, such as by network traffic
analysis, and state our adversary model.

@@

\paragraph*{Adversary model} To assess our approach, we use the
adversary model assumed by \ac{Tor} and Sphinx~\cite{Sphinx}:
the adversary has complete control over some part of the network as
well as some of the nodes. Like them, we do not consider a passive global network
adversary, against which it is notoriously difficult to achieve
security~\cite{SystemsForAnonymousCommunication}.
